Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2021-01-02 xrb2/paper.tex
added more sec 3.1 text (in blue)

@@

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
increasing the 
neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
gradient becomes important, and is approximately:
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. Comparing Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, the horizontal extent of the $\tt{F1000}$ flame does appear to be 
reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
here the difference is not a factor of two. The simulations seem to follow the theoretical
scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
up), based on the results of \citet{flame_wave1}.}

The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in 
\citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel 
layer --- but they differ in their burning structures. Figures~\ref{fig:time_series_500} and 
\ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt 
F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not 
reach as high atomic weights. This is not surprising, since those previous runs artificially 
boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the 
higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the 
Figure  \ref{fig:time_series_1000} 100 ms snapshot). We believe that this is because the increased 
rotation better confines the initial perturbation and subsequent expansion from the burning, 
increasing the temperature and density in the flame front such that the reaction rate increases, 
which allows the reactions to progress further. In contrast to {\tt F500} and {\tt F1000}, the 
lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 
aligns with the reasoning given above, with the lower rotation in this case potentially leading to 
insufficient confinement such that the temperature and density required for ignition is not 
achieved. In this scenario, another source of confinement (e.g.\ magnetic fields) would need to 
take over at lower rotation rates to allow a burst to occur, at least for the initial model used 
here.

>>>>

We run three models ({\tt F250}, {\tt F500}, and {\tt F1000}) with the same initial model in terms 
of temperature but differing rotation rates. {\color{blue}We saw in \citet{flame_wave1} that 
increasing the 
neutron star rotation rate reduces the horizontal lengthscale of the flame. An estimate of 
this lengthscale is given by the Rossby radius of deformation, $L_R$. The Rossby radius may be 
thought of as the scale over which the balance between the Coriolis force and horizontal pressure 
gradient becomes important, and is approximately:
\begin{equation}
	\label{eqn:rossby}
	L_R \approx \frac{\sqrt{g H_0}}{\Omega},
\end{equation}
where $g$ is gravitational acceleration, $H_0$ is the atmospheric scale height, and $\Omega$ is the 
neutron star rotation rate. In Figure \ref{fig:time_series_enuc_500} and Figure 
\ref{fig:time_series_enuc_1000}, we use the energy generation rate $\dot{e}_\mathrm{nuc}$ at $50~\mathrm{ms}$ and $100~\mathrm{ms}$ for {\tt F500} and {\tt F1000} to discern the horizontal extent of the flame at different rotation rates. Taking the rightmost edge of the bright teal/green region (where significant energy generation is occuring) as the leading edge of the flame in each plot, we see that the horizontal extent of the $\tt{F1000}$ flame appears to be 
reduced compared to the lower rotation $\tt{F500}$ run. From Equation \ref{eqn:rossby}, increasing 
the rotation rate from 500 Hz to 1000 Hz should decrease $L_R$ by a factor of two, and the greater 
confinement from the Coriolis force should reduce the horizontal extent of the flame by a similar 
factor. The Rossby radius is not a perfect measure of this horizontal lengthscale, however, and 
here the difference is not a factor of two. The simulations seem to follow the theoretical
scaling described in Equation \ref{eqn:rossby} more closely at higher rotation rates (1000 Hz and 
up), based on the results of \citet{flame_wave1}.}

@@

The {\tt F500} and {\tt F1000} runs both qualitatively resemble the flame structure in 
\citet{flame_wave1} --- a laterally propagating flame that is lifted off of the bottom of the fuel 
layer --- but they differ in their burning structures. Figures~\ref{fig:time_series_500} and 
\ref{fig:time_series_1000} show time series of the mean molecular weight, $\bar{A}$, for the {\tt 
F500} and {\tt F1000} runs.  Compared to those in \cite{flame_wave1}, ashes behind the flame do not 
reach as high atomic weights. This is not surprising, since those previous runs artificially 
boosted the reaction rates.  Comparing these two new runs, the burning is much more evolved for the 
higher rotation rate, and the ash is actually able to move ahead of the flame front (visible in the 
Figure  \ref{fig:time_series_1000} 100 ms snapshot). We believe that this is because the increased 
rotation better confines the initial perturbation and subsequent expansion from the burning, 
increasing the temperature and density in the flame front such that the reaction rate increases, 
which allows the reactions to progress further. {\color{blue} The $\dot{e}_\mathrm{nuc}$ plots in Figure \ref{fig:time_series_enuc_1000} also support this interpretation, with the interior of the {\tt F1000} flame reaching higher $\dot{e}_\mathrm{nuc}$ values than for the {\tt F500} run in Figure \ref{fig:time_series_enuc_500}.} In contrast to {\tt F500} and {\tt F1000}, the 
lowest rotation run --- {\tt F250} --- failed to ignite. The lack of ignition for {\tt F250} also 
aligns with the reasoning given above, with the lower rotation in this case potentially leading to 
insufficient confinement such that the temperature and density required for ignition is not 
achieved. In this scenario, another source of confinement (e.g.\ magnetic fields) would need to 
take over at lower rotation rates to allow a burst to occur, at least for the initial model used 
here.
